---
title: "Demand Forecasting for a Fast-Food Restaurant Chain"
author: "Konstantinos Paganopoulos"
subtitle: Logistics and Supply Chain Analytics - Individual Project
output:
  pdf_document: default
  html_document:
    df_print: paged
---

### Solutions

We first load the necessary libraries.

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
library(forecast)
library(tseries)
library(ggplot2)
```

We have a dataset, which includes daily sales for lettuce at a store in New York from a fast-food restaurant chain from 20 March 2015 to 15 June 2015. Each observation includes two values: day pair, and sales in that particular day.

Then, we load and split the data set into train and test set.

```{r}
# read csv file
data <- read.csv(file = "NewYork2_final.csv", header = TRUE, stringsAsFactors = FALSE)

# convert column date of data set to type date
data$date <- as.Date(data$date)

# convert sales into a time series object
lettuce <- ts(data[, 2], frequency = 7, start = c(12, 2)) # 12th week 2nd day

# split data set into train and test set
lettuce_train <- subset(lettuce, end = 80, start = 7) # ignore first 6 days due to strange data
lettuce_test <- subset(lettuce, start = 81) # last 14 lines-days 
```

### ARIMA

We visually inspect the time series.

```{r}
autoplot(lettuce_train, xlab = "Weeks", ylab = "Lettuce") + theme_minimal() + 
ggtitle("New York 2 Store (20974) - Time series plot")
```
```{r}
ggtsdisplay(lettuce_train, xlab = "weeks", theme = theme_minimal())
```

We plot the time series, and observe that there is no seasonality and appears to be stationary. We run ADF, PP and KPSS tests to formally test the stationarity of time series and all suggest that the time series is stationary.

```{r}
# stationary test
adf.test(lettuce_train)
```

```{r}
pp.test(lettuce_train)
```

```{r}
kpss.test(lettuce_train)
```

The two automatic functions, ndiffs() and nsdiffs() tell us how many first-order differences, and how many seasonal differences, respectively, we need to take to make the time series stationary. We use those functions below:

```{r}
ndiffs(lettuce_train)
```
```{r}
# seasonal stationarity
nsdiffs(lettuce_train)
```

We do not need to differentiate.

We plot again the time series to see if stationary.

```{r}
autoplot(lettuce_train, xlab = "Weeks", ylab = "Lettuce") + theme_minimal() + 
ggtitle("New York 2 Store (20974) - Time series plot")
```

We can easily observe a concave pattern. This forms a trend.

As a result, we have to differentiate as follows:

```{r}
lettuce_train.diff1 <- diff(lettuce_train, differences = 1)
```

We plot again the time series to see if stationary.

```{r}
autoplot(lettuce_train.diff1, xlab = "Weeks", ylab = "Lettuce") + theme_minimal() + 
ggtitle("New York 2 Store (20974) - Time series plot")
```

Looks stationary.

The next step is to determine the optimal orders of MA and AR components. We first plot the ACF and PACF of the time series.

```{r}
# acf plot
ggAcf(lettuce_train.diff1) + theme_minimal() + ggtitle("New York 2 Store (20974) - ACF plot")
```
```{r}
# pacf plot
ggPacf(lettuce_train.diff1) + theme_minimal() + ggtitle("New York 2 Store (20974) - PACF plot")
```

Next we use $auto.arima()$ to search for the best ARIMA models. 

The default procedure uses some approximations to speed up the search. These approximations can be avoided with the argument approximation = FALSE. It is possible that the minimum AIC model will not be found due to these approximations, or because of the stepwise procedure. A much larger set of models will be searched if the argument stepwise = FALSE is used. We also use d = 1 and D = 0 since we had only first differenting.

```{r}
auto.arima(lettuce_train, trace = TRUE, ic = 'aic', approximation = FALSE, stepwise = FALSE, d=1, D=0) 
# Best model: ARIMA(1,1,1)(1,0,0)[7]  (AIC=783.51)
# Second best:  ARIMA(0,1,2)(1,0,0)[7]  (AIC=783.85)
# Third best: ARIMA(0,1,1)(1,0,0)[7] (AIC=784.06)
```

Based on the output of $auto.arima()$, a couple of models have similar AICs. Now suppose that we choose the three models with the lowest AICs, namely ARIMA(1,1,1)(1,0,0)[7] with AIC=783.51, ARIMA(0,1,2)(1,0,0)[7]  with AIC=783.85 AND ARIMA(0,1,1)(1,0,0)[7] with AIC=784.06, as the candidate models that we would like to evaluate further.

```{r}
# three candidate models
lettuce.m1 <- Arima(lettuce_train, order = c(1, 1, 1), 
                       seasonal = list(order = c(1, 0, 0), period = 7))
lettuce.m2 <- Arima(lettuce_train, order = c(0, 1, 2), 
                       seasonal = list(order = c(1, 0, 0), period = 7))
lettuce.m3 <- Arima(lettuce_train, order = c(0, 1, 1), 
                       seasonal = list(order = c(1, 0, 0), period = 7))
```

Now we evaluate the in-sample performance/fit of the model with $accuracy()$ function, which summarizes various measures of fitting errors. 

A couple of functions are proved to be useful for us to evaluate the in-sample performance/fit of the model. One is accuracy() function, which summarizes various measures of fitting errors. In the post-estimation analysis, we would also like to check out the residual plots, including time series, ACFs and etc, to make sure that there is no warning signal. In particular, residuals shall have a zero mean, constant variance, and distributed symmetrically around mean zero. ACF of any lag greater 0 is expected to be statistically insignificant.

```{r}
# in-sample one-step forecasts model 1
accuracy(lettuce.m1)
```
```{r}
# in-sample one-step forecasts model 2
accuracy(lettuce.m2)
```

```{r}
# in-sample one-step forecasts model 3
accuracy(lettuce.m3)
```

The first model even though it has both the lowest AIC score as well as the lowest RMSE.

Now we proceed with the residual analysis of the three models.

```{r}
# residual analysis model 1
autoplot(lettuce.m1$residuals, xlab = "Weeks", ylab = "Lettuce") + theme_minimal() + 
ggtitle("New York 2 Store (20974) - Residuals model 1 plot")
```
```{r}
ggAcf(lettuce.m1$residuals) + theme_minimal() + 
ggtitle("New York 2 Store (20974) - ACF residualts plot model 1")
```
```{r}
checkresiduals(lettuce.m1, xlab = "weeks", theme = theme_minimal())
```
```{r}
# residual analysis model 2
autoplot(lettuce.m2$residuals, xlab = "Weeks", ylab = "Lettuce") + theme_minimal() + 
ggtitle("New York 2 Store (20974) - Residuals model 2 plot")
```
```{r}
ggAcf(lettuce.m2$residuals) + theme_minimal() + 
ggtitle("New York 2 Store (20974) - ACF residualts plot model 2")
```
```{r}
checkresiduals(lettuce.m2, xlab = "weeks", theme = theme_minimal())
```

```{r}
# residual analysis model 3
autoplot(lettuce.m3$residuals, xlab = "Weeks", ylab = "Lettuce") + theme_minimal() + 
ggtitle("New York 2 Store (20974) - Residuals model 3 plot")
```

```{r}
ggAcf(lettuce.m3$residuals) + theme_minimal() + 
ggtitle("New York 2 Store (20974) - ACF residualts plot model 3")
```

```{r}
checkresiduals(lettuce.m3, xlab = "weeks", theme = theme_minimal())
```

Now we continue with the forecasting part for the three candidate models:

```{r}
#Forecasting part model 1
lettuce.f1 <- forecast(lettuce.m1, h = 14)
autoplot(lettuce.f1, xlab = "Weeks", ylab = "Lettuce") + theme_minimal()
```

```{r}
#Forecasting part model 2
lettuce.f2 <- forecast(lettuce.m2, h = 14)
autoplot(lettuce.f2, xlab = "Weeks", ylab = "Lettuce") + theme_minimal()
```

```{r}
#Forecasting part model 3
lettuce.f3 <- forecast(lettuce.m3, h = 14)
autoplot(lettuce.f3, xlab = "Weeks", ylab = "Lettuce") + theme_minimal()
```

Now we need to test how our models performs for test set. Earlier observations are used for training, and more recent observations are used for testing. Suppose we use the first 80 days of data for training and the last 14 for test. Based on auto.arima(), we choose two candidate models with the lowest AICs.

```{r}
### model evaluation
# Apply fitted model to later data
# Accuracy test for candidate model 1
accuracy.m1 <- accuracy(forecast(lettuce.m1, h = 14), lettuce_test)
accuracy.m1
```
```{r}
# Accuracy test for candidate model 2
accuracy.m2 <- accuracy(forecast(lettuce.m2, h = 14), lettuce_test)
accuracy.m2
```
```{r}
# Accuracy test for candidate model 3
accuracy.m3 <- accuracy(forecast(lettuce.m3, h = 14), lettuce_test)
accuracy.m3
```

Thus we pick the first model, since it performs better on the test set.

Now we train the first model on the whole date set as follows:

```{r}
# Training on both train and test set
lettuce.f.both <- Arima(lettuce, order = c(1, 1, 1), 
                    seasonal = list(order = c(1, 0, 0), period = 7))
```

Lastly, we forecast lettuce demand for the next 2 weeks.

```{r}
# Forecast for next 14 days
lettuce.f.final <- forecast(lettuce.f.both, h = 14)
lettuce.f.final
```

We present our forecast through ARIMA(1,1,1)(1,0,0) model for each of the next 14 days.

```{r}
forecast_data <- as.data.frame(lettuce.f.final)
next2weeks <- data.frame(day = seq(1, 14))
final_forecast_NewYork2_arima <- cbind(next2weeks, forecast_data$`Point Forecast`)
final_forecast_NewYork2_arima
```

### Holt-Winters 

Now we will use another model to forecast lettuce demand. Our goal is to pick the model with the most accurate predictions.

We will forecast the lettuce demand for next two weeks using Holt-Winters model.

For time series analysis, the first step is always to visually inspect the time series. In this regard, the stl() function is quite useful. It decomposes the original time series into trend, seasonal factors, and random error terms. The relative importance of different components are indicated by the grey bars in the plots.

```{r}
lettuce_train %>% stl(s.window = "period") %>%
autoplot(xlab = "Weeks", ylab = "Lettuce") + theme_minimal() + 
ggtitle("New York 2 Store (20974) - Range bar plot")
```

For this data set, the grey bar of the trend panel is significantly larger than that on the orginal time series panel, which indicates that the contribution of the trend component to the variation in the original time series is marginal.

The grey bar of the seasonal panel is also large, and larger than the grey bar of random error term, which indicates that the contribution of seasonal factors to the variation in the original time series is marginal too. In other words, it indicates that there is no seasonality in the data.

With ets(), initial states and smoothing parameters are jointly estimated by maximizing the likelihood function. We need to specify the model in ets() using three letters. The way to approach this is: (1) check out time series plot, and see if there is any trend and seasonality; (2) run ets() with model = “ZZZ”, and see whether the best model is consistent with your expectation; (3) if they are consistent, it gives us confidence that our model specification is correct; otherwise try to figure out why there is a discrepancy.

We now use ets function as previously indicated to find our best model:

```{r}
# using ets
lettuce.ets2 <- ets(lettuce_train, model = "ZZZ")
lettuce.ets2
```

Our best model is the ETS(A,N,A).

```{r}
# using ets
lettuce.ets <- ets(lettuce_train, model = "ANA", ic = 'aic')
lettuce.ets
```

After estimation, we can use accuracy() function to determine in-sample fit and forecast() function to generate forecast. 

Similarly with ARIMA model, we use AIC to determine our best model in terms of best in-sample performance.

```{r}
# in-sample one-step forecast
accuracy(lettuce.ets)
```

We present the in-sample forecast part for the ets model as follows:

```{r}
# best model
lettuce.ets.f <- forecast(lettuce.ets, h = 14)
lettuce.ets.f
```

After the forecast, we continue with the in and out of sample accuracy of the two ets models.

```{r}
# Out of sample accuracy
# best model
accuracy.ets <- accuracy(lettuce.ets.f, lettuce_test)
accuracy.ets
```

We now train our best model - ETS(A,N,A) on the whole data set as indicated below:

```{r}
# final model
lettuce.ets <- ets(lettuce, model = "ANA", ic = 'aic')
lettuce.ets
```

We now present the out-of-sample forecast for the next 14 days (2 weeks) as seen below:

```{r}
lettuce.ets.f <- forecast(lettuce.ets, h = 14)
lettuce.ets.f
```

We present our forecast for each of the next 14 days.

```{r}
forecast_data <- as.data.frame(lettuce.ets.f)
next2weeks <- data.frame(day = seq(1, 14))
final_forecast_NewYork2_ets <- cbind(next2weeks, forecast_data$`Point Forecast`)
final_forecast_NewYork2_ets
```

### Comparison

Now we will compare the two best models for New York 2 Store (20974).

We plot time series data for train and test set and also the forecasts from our two models as indicated below:

```{r}
colours <- c("blue", "deepskyblue4", "black")
autoplot(lettuce.f.final, xlab = "Weeks", ylab = "Lettuce") + 
  autolayer(lettuce_train, series = "Train set") +
  autolayer(lettuce_test, series = "Test set") +
  autolayer(lettuce.f.final, series = "14 day Forecast") +
  guides(colour = guide_legend(title = "Time Series Data")) +
  scale_colour_manual(values = colours) + theme_minimal()
```
```{r}
autoplot(lettuce.ets.f, xlab = "Weeks", ylab = "Lettuce") + 
  autolayer(lettuce_train, series = "Train set") +
  autolayer(lettuce_test, series = "Test set") +
  autolayer(lettuce.ets.f, series = "14 day Forecast") +
  guides(colour = guide_legend(title = "Time Series Data")) +
  scale_colour_manual(values = colours) + theme_minimal()
```

In order to decide which of the two models ARIMA(0,1,2)(1,0,0) or ETS(A,N,A) to choose, we will check their RMSE in the test set.

```{r}
# best ets model
# ETS(A,N,A) 
accuracy.ets
```
```{r}
# best arima model
# ARIMA(1,1,1)(1,0,0)
accuracy.m2
```

We can observe that ARIMA(1,1,1)(1,0,0) has a better (lower) RMSE (41.21143 vs 55.99135) respectively.

Therefore, we choose the ARIMA(1,1,1)(1,0,0) for New York (20974) store.

Hence, our forecast for lettuce demand of next 2 weeks for that store is the following:

```{r}
final_forecast_NewYork2_arima
```
